Illustration of Adaptive Shrinkage

### 2016-10-03 analysis/adaptive_shrinkage.Rmd
Made a few slight adjustments to adaptive_shrinkage.Rmd, mainly formatting of the code. It works with 'general' branch of ashr.

```{r plot_shrunk_vs_obs}
make_df_for_ashplot =
  function (sim1, sim2, ash1, ash2, name1 = "spiky", name2 = "big-normal") {
    n    = length(sim1$input$betahat)
    lfsr = c(get_lfsr(ash1),get_lfsr(ash2))
    return(data.frame(betahat  = c(sim1$input$betahat,sim2$input$betahat),
                      beta_est = c(get_pm(ash1), get_pm(ash2)),
                      lfsr     = lfsr,
                      s        = c(sim1$input$sebetahat,sim2$input$sebetahat),
                      scenario = c(rep(name1,n),rep(name2,n)),
                      signif   = lfsr < 0.05))
  }

ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate") {
  ggplot(df,aes(x = betahat,y = beta_est,color = 1/s)) +
    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +   geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
    scale_colour_gradient2(midpoint = median(1/s),low = "blue",mid = "white",
                           high = "red",space = "Lab") +
    coord_fixed(ratio = 1)
 }

@@

z_lfsr_plot=function(df,ylab="Observed Z score",xlab="lfsr"){
  ggplot(df,aes(x=lfsr,y=betahat/s,color=1/s)) +
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) + scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")
}

@@

```{r plot_pvalues}
pval_plot=function(df){
  ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
  scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")
}
pval_plot(df)
```

Let's see how these are affected by changing the modelling assumptions so that
the standardized beta are exchangeable (rather than the beta being exchangeable).
```{r run_ash_ET}
beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")
beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,alpha=1,mixcompdist="normal")
df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)
ashplot(df.ET,ylab="Shrunken beta estimate (ET model)")
pval_plot(df.ET)
```

This is a "volcano plot" showing effect size against p value. The blue points are "significant" in that they have lfsr<0.05. In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
```{r volcano}
  ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
  xlim(c(-10, 10)) + ylim(c(0, 15)) +
  xlab("Effect (beta)") + ylab("-log10 p-value")
```

```{r sehist}
hist(s,main="histogram of standard errors")
summary(s)
```

## Session information

```{r info}
sessionInfo()
```

>>>>

```{r plot_shrunk_vs_obs}
make_df_for_ashplot =
  function (sim1, sim2, ash1, ash2, name1 = "spiky", name2 = "big-normal") {
    n = length(sim1$input$betahat)
    x = c(get_lfsr(ash1),get_lfsr(ash2))
    return(data.frame(betahat  = c(sim1$input$betahat,sim2$input$betahat),
                      beta_est = c(get_pm(ash1),get_pm(ash2)),
                      lfsr     = x,
                      s        = c(sim1$input$sebetahat,sim2$input$sebetahat),
                      scenario = c(rep(name1,n),rep(name2,n)),
                      signif   = x < 0.05))
  }

ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate")
  ggplot(df,aes(x = betahat,y = beta_est,color = 1/s)) +
    xlab(xlab) + ylab(ylab) + geom_point() +
    facet_grid(.~scenario) +
    geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
    scale_colour_gradient2(midpoint = median(1/s),low = "blue",
                           mid = "white",high = "red",space = "Lab") +
    coord_fixed(ratio = 1)

@@

z_lfsr_plot = function(df,ylab = "Observed Z score",xlab = "lfsr")
  ggplot(df,aes(x = lfsr,y = betahat/s,color = 1/s)) +
    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +
    scale_colour_gradient2(midpoint = median(1/s),low = "blue",
                           mid = "white",high = "red",space = "Lab")

@@

**Note:** I find it confusing that you are using the same colour scheme here as
above, but you are showing `log(s)`, whereas above you are showing `1/s`. It
seems that it would be better to reverse the colour scheme for the `pval_plot`
to better correspond to the above plots. –Peter

```{r plot_pvalues}
pval_plot = function (df)
  ggplot(df,aes(x = pnorm(-abs(betahat/s)),y = lfsr,color = log(s))) +
  geom_point() + facet_grid(.~scenario) + xlim(c(0,0.025)) +
  xlab("p value") + ylab("lfsr") +
  scale_colour_gradient2(midpoint = 0,low = "blue",
                         mid = "white",high = "red")

print(pval_plot(df))
```

Let's see how these are affected by changing the modelling assumptions so that
the *standardized* beta are exchangeable (rather than the beta being exchangeable).

```{r run_ash_ET}
beta.bignormal.ash.ET =
  ash(sim.bignormal$input$betahat,s,alpha = 1,mixcompdist = "normal")
beta.spiky.ash.ET =
  ash(sim.spiky$input$betahat,s,alpha = 1,mixcompdist = "normal")
df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,
                            beta.bignormal.ash.ET)
ashplot(df.ET,ylab = "Shrunken beta estimate (ET model)")
pval_plot(df.ET)
```

This is a "volcano plot" showing effect size against p value. The blue points
are "significant" in that they have lfsr < 0.05.

```{r volcano}
print(ggplot(df,aes(x = betahat,y = -log10(2*pnorm(-abs(betahat/s))),
                    col = signif)) +
  geom_point(alpha = 1,size = 1.75) + facet_grid(.~scenario) +
  theme(legend.position = "none") + xlim(c(-10,10)) + ylim(c(0,15)) +
  xlab("Effect (beta)") + ylab("-log10 p-value"))
```

In this case the significance by lfsr is not quite the same as cutting off
at a given p value (you can see that the decision boundary is not quite the
same as drawing a horizontal line), but also not that different, presumably
because the standard errors, although varying across observations, do not
vary greatly.

```{r sehist}
hist(s,main = "histogram of standard errors")
print(summary(s))
```

Session information.

```{r info}
print(sessionInfo())
```
